---
title: "p8105_hw2_sl4657"
author: "Si Li"
output: github_document
editor_options: 
  chunk_output_type: console
---

# Problem 1
## Use Mr.Trash.Wheel sheet
```{r}
library(tidyverse)
library(readxl)
library(dplyr)
mtw_data = 
  read_excel(
    "data files/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet=1) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster))%>% 
  select(-15) %>% 
  mutate(sports_balls = round(sports_balls, digits = 0)) %>% 
  mutate(sports_balls = as.integer(sports_balls))
mtw_data
```

```{r}
p8_data = 
  readxl::read_excel(
    "./data files/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet=5,range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year= "2018")
p8_data

p7_data = 
  readxl::read_excel(
    "./data files/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet=6,range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year= "2017")
p7_data

p78_data=
  bind_rows(p7_data, p8_data) %>% 
  mutate(month = month.name[month]) %>% 
  select(year, everything())
p78_data
```
The number of observations in 2018 precipitation dataset is `r nrow(p8_data)`, in 2017 precipitation dataset is `r nrow(p7_data)`, key variables are `r colnames(p78_data[1:2])`, which are used to combine 2018 precipitation dataset and 2017 precipitation dataset. The total precipitation in 2018 is `r sum(pull(p8_data, total))`, the median number of sports balls in a dumpster in 2017 is `r median(pull(mtw_data,sports_balls), 167:221)`.

# Problem 2
##  Use the FiveThirtyEight data
```{r}
pols_data = 
  read_csv(file = "./data files/pols-month.csv") %>% 
  separate(col = mon, into = c("year", "month", "day"), "-") %>%
  mutate(
    month = as.numeric(month),
    month = month.abb[month])%>% 
  mutate(president = recode(prez_gop, `0` ="dem", `1` = "gop_1", `2`= "gop_2")) %>% 
  select(-day,- prez_gop, -prez_dem)
pols_data

snp_data =
  read_csv(file = "./data files/snp.csv") %>% 
  separate(col = date, into = c("month", "day", "year"), "/") %>% 
  mutate(month = as.numeric(month))%>% 
  arrange(year, month) %>% 
  mutate(month=month.abb[month]) %>% 
  select(year, month, close) 
snp_data

unem_data =
  read_csv(file = "./data files/unemployment.csv") %>% 
  pivot_longer(
    Jan: Dec,
    names_to = "month",
    values_to = "unem_perc"
  ) %>% 
  janitor::clean_names()
unem_data

s_p = 
  full_join(pols_data, snp_data, by= c("year","month")) %>% 
  mutate(year= as.double(year))

s_p_u=
  full_join(s_p, unem_data, by= c("year", "month"))
s_p_u

```
In the pols_data, dimension is `r dim(pols_data)`, range of years is `r range(pull(pols_data, year))`,  key variables are `r colnames(pols_data[1:2])`.
In the snp_data, dimension is `r dim(snp_data)`, range of years is `r range(pull(snp_data, year))`,  key variables are `r colnames(snp_data[1:2])`.
In the unem_data, dimension is `r dim(unem_data)`, range of years is `r range(pull(unem_data, year))`,  key variables are `r colnames(unem_data[1:2])`.
Key variables in order to combine these datasets are the same.

# Problem 3
## Use data on Popularity of baby names
```{r}
pbn_data=
  read_csv(file= "./data files/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(gender= str_to_lower(gender),
         ethnicity= str_to_lower(ethnicity),
         childs_first_name = str_to_title(childs_first_name)) %>% 
  distinct(year_of_birth, gender, ethnicity, childs_first_name, .keep_all = TRUE)
pbn_data

oli_rank=
  pbn_data %>%
  filter(childs_first_name %in% "Olivia")
oli_rank

eth_rank=
  pbn_data %>% 
  filter(childs_first_name %in% "Ethan")
eth_rank

pbn_data %>% 
  filter(gender=="male",
         ethnicity=="white non hispanic",
         year_of_birth=="2016") %>% 
  ggplot(aes(x=rank, y=count))+
  geom_point()+
  labs(title="Relationship between number and rank in popularity of the name of a male, white non-hispanic child born in 2016")
```

